Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-04-04 paper.Rnw
Fixed QC results, and added a table of segmentation totals for ADNI1 Screening Dataset

\newcommand{\mb}{MAGeT Brain}

@@

Multi-atlas methods have been very successfully applied to hippocampal
segmentation. Collins et al. found near-manual segmentation performance using an
atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
ANIMAL nonlinear registration algorithm, normalised mutual information as an
atlas selection similarity metric, and majority vote for label fusion
\citep{Collins2010}.  The Alzheimer's Disease Neuroimaging Initiative (ADNI) is
a commonly used benchmarking dataset of MR images of controls and patients with
MCI or Alzheimer's (see Methods for more information on the ADNI dataset).
Leung et al. tuned parameters for registration and label fusion to the
segmentation of ADNI1 1 year dataset of images with an atlas library of 55
images \citep{Leung2010}.  The MAPER whole brain segmentation algorithm
\citep{Heckemann2006,Heckemann2011}, using 30 atlases, on all ADNI1 baseline
images \citep{Heckemann2011}.  Lotjonen et al. use images from the ADNI1
baseline dataset and 30 atlases with a proprietary non-linear registration
method based on intensity differences, and post-processing step using a graph
cuts algorithm to optimise fused segmentations against a spatial intensity prior
\citep{Lotjonen2012}.  
\todo{No story here}

@@

\subsubsection{Automatic Normalization and Image Matching and Anatomical
Labeling (ANIMAL)}

The ANIMAL algorithm carries out image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the whole
brain \citep{Collins}. In the second phase, nonlinear registration is completed
using the ANIMAL algorithm \citep{Collins1995}: an iterative procedure that
estimates a 3D deformation field between two MR images. At first, large
deformations are estimated using blurred version of the input data. These larger
deformations are then input to subsequent steps where the fit is refined by
estimating smaller deformations on data blurred with a Gaussian kernel with a
smaller FWHM. The final transformation is a set of local translations defined on
a bed of equally spaced nodes that were estimated through the optimization of
the correlation coefficient.  For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004},
displayed in table \ref{tab:ANIMAL_params}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANIMAL parameters                                                           %
%                                                                             %
\begin{table*}[!tbp]
\scriptsize
\caption{ANIMAL Registration Parameters\label{tab:ANIMAL_params}} 
\begin{center}
\begin{tabular}{l | c c c}
\hline 
Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
\hline 
Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Iterations        &   30    &    30   &   10     \tabularnewline
Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
Sub-Lattice       &   6     &    6    &   6      \tabularnewline
Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
\hline
\end{tabular}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Automatic Normalization Tools (ANTS)}

ANTs is a diffeomorphic registration algorithm which provides great flexibility
over the choice of transformation model, objective function, and the consistency
of the final transformation. The transformation is estimated in a hierarchical
fashion where the MRI data is subsampled, allowing large deformations to be
estimated and successively refined at later hierarchical stages (where the data
is subsampled to a finer grid). The deformation field and the objective function
are regularized with a Gaussian kernel at each level of the hierarchy. The ANTs
algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
implementation of the ANTS algorithm compatible with the MINC data format,
mincANTS \url{https://github.com/vfonov/mincANTS}.

We used the following command line when running ANTS:
{\small
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
}
These settings were adapted from the "reasonable starting point" given in the
ANTS manual 
\footnote{\url{https://sourceforge.net/projects/advants/files/Documentation/}}.

@@

The {\tt xcorr\_vol} utility from the ANIMAL toolkit is used to calculate the
cross-correlation similarity measure.

@@

%%%%%
%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
%%%%%
\subsubsection{ADNI1 cross-validation}
To test the effect of parameter settings on the \mb algorithm, Monte
Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
the ADNI1 subjects for validation.  This form of cross-validation (described in
detail below) allows us to validate \mb with various atlas and template
library sizes, registration and label fusion methods.

@@

{\bf Registration method.} Both the ANTS and ANIMAL registration methods were
used.

@@

A total of 10 validation rounds are perfomed on each subject in the dataset, for each 
combination of parameter settings: atlas library size (1-9), template library 
size (1-20), registration method (ANTS vs. ANIMAL), and label fusion method 
(MV, XC-WV, NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were
conducted, resulting in a total of 1'490'400 segmentations analysed.

@@

{\bf Registration method.} The ANTS non-linear registration algorithm was used,
as it performed best in the ADNI1 cross-validation experiment (see Results).

@@

{\bf Evaluation.}
Since the hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct similarity comparisons
between labels produced by \mb and the ADNI labels.  To evaluate the
performance of \mb, we compare the correlation of \mb segmentation
volumes with manual segmentation (SNT) volumes. Additionally, we correlate the
hippocampal volumes of established automated segmentation methods to \mb
segmentations.

Additionally, we compared classification accuracy of subjects by diagnosis based
on hippocampal volume using both the SNT labels and our produced labels.

\todo{explain why we did(n't) resegment the ADNI images with the low-res
protocol and compare directly like that.}

\todo{Also: group differences, expecting separation}

@@

{\bf Registration method.} ANTS.
\todo{complete}

@@

{\bf Registration method.} ANTS
\todo{complete}

@@

%%%%%
%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
%%%%%
\subsection{ADNI-1 Cross-Validation}

@@

In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of 
each of 69 subjects, over a range of atlas and template library sizes, with ANTS 
and ANIMAL registration algorithms, and three kinds of label fusion techniques. 
Hippocampal \mb-based segmentations using both ANIMAL and ANTs registration
algorithm demonstrate good overlap with manually derived gold-standards (Figure
\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both ANIMAL and ANTs-based 
segmentations demonstrate trend overlap accuracy that increases with the size of 
atlas library and template library. With 12 or more templates \mb library of at 
least 12 templates

No marked difference in segmentation accuracy is seen when either ANIMAL or ANTs
registration is used with different label fusion techniques, at any atlas or
template library sizes.  In every parameter configuration, the use of \mb with 
ANTs registration shows increased segmentation accuracy over \mb with ANIMAL registration (\todo{interaction?}).

It is interesting to note that with an even number of templates, \mb shows a
small decrease in performance relative to when one fewer template image is used
(\todo{show significance of this}). See section (\todo{ref}) for a discussion of
this behaviour. In the remainder of this section, only results from odd-sized
template libraries will be shown.

@@

ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" )
@
  \caption{\mb mean DSC by atlas and template library size}
  \label{fig:ADNI1_xval-k_mean}
\end{figure}

With an increasing number of templates, \mb-based segmentating using ANTs
registration and majority vote label fusion shows improvement in overlap
accuracy over multi-atlas-based segmentation, using the same number of atlases
and voting method Figure \todo{ref}. The magnitude of improvement over
multi-atlas-based segmentation decreases with an increasing number of atlases,
with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (0.02
DSC) is found when one atlas is used with a template library of 20 images.
\todo{stats?}

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  coord_cartesian(ylim=c(-0.1,0.05)) + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
@
  \caption{Increase in mean Kappa between \mb and multi-atlas}
  \label{ADNI1_cross_validation_MA_difference}
\end{figure}

@@

  #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
  #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
  #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
  #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
  labs(x="Number of Atlases", y = "Kappa", 
       title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
  scale_colour_hue(name="Method") + 
  coord_cartesian(ylim=c(0.7,0.9))
  #theme(legend.justification=c(1,0), legend.position=c(1,0))
@
  \caption{Change in mean Kappa between multi-atlas and \mb using ANTS,
  20 templates}
  \label{}
\end{figure}

@@

%%%%%
%%%%%                  Results: ADNI-1 Screening Validation               %%%%%
%%%%%
\subsection{ADNI-1 Screening Validation}

@@

<<ADNI-src-volumes-prep,echo=F,cache=T>>=
#bl1.5T= read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
#bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
ucsd  = read.csv("data/ADNI_baseline_volumes/UCSDVOL.csv")
fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
# recode VISCODE
#viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
#yr1$VISCODE <- as.factor(yr1$Visit)
#levels(yr1$VISCODE) <-viscode_levels
#bl1.5T$VISCODE <- 'bl'
#bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)

#TODO: include Scaled_2? 
subjects = subset(yr1,Visit ==1)
subjects$VISCODE = 'bl'

# fetch just the columns we need, and do a little renaming
mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, 
                           MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, 
                           MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, 
                           SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
ucsd_pruned   = data.frame(RID=ucsd$RID, VISCODE=ucsd$VISCODE, 
                           UCSD_L=ucsd$LHIPPOC, UCSD_R=ucsd$RHIPPOC)
fs_pruned     = subset(fs, TEMPQC != "Fail")
fs_pruned     = data.frame(FS_L=fs_pruned$ST29SV, 
                           FS_R=fs_pruned$ST88SV, Source = fs_pruned$IMAGEUID)
fsl_pruned    = data.frame(FSL_L=fsl$X17, 
                           FSL_R=fsl$X53, Source=fsl$Source)
qc_pruned     = qc[,c("Image.Data.ID","MAGeT_QC","FSL_QC")]
qc_pruned     = subset(qc_pruned, MAGeT_QC == 1 & FSL_QC == 1)

# diagnoses
dx = adnimerge[c("DX.bl", "RID", "VISCODE")]

# Now create one data.frame with all the yr1 volume data we have from each datasource
combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
#combined = merge(combined, ucsd_pruned , by=c("RID", "VISCODE"), all.x =TRUE)              # TODO: this a unique key?
combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
combined = merge(combined, qc_pruned   , by=c("Image.Data.ID"))

attach(combined)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  #UCSD  = UCSD_L + UCSD_R,
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)

detach(combined)
#complete_totals = subset(totals, !is.na(SNT) & !is.na(FSL) & !is.na(FS) & !is.na(MAPER) & !is.na(MAGeT))
totals.complete = na.omit(totals)
@

>>>>

\newcommand{\mb}{MAGeT Brain}
\newcommand{\ants}{ANTS}
\newcommand{\animal}{ANIMAL}

@@

Multi-atlas methods have been very successfully applied to hippocampal
segmentation. Collins et al. found near-manual segmentation performance using an
atlas library of 80 T1-weighted atlas images from the ICBM152 dataset, the
\animal nonlinear registration algorithm, normalised mutual information as an
atlas selection similarity metric, and majority vote for label fusion
\citep{Collins2010}.  The Alzheimer's Disease Neuroimaging Initiative (ADNI) is
a commonly used benchmarking dataset of MR images of controls and patients with
MCI or Alzheimer's (see Methods for more information on the ADNI dataset).
Leung et al. tuned parameters for registration and label fusion to the
segmentation of ADNI1 1 year dataset of images with an atlas library of 55
images \citep{Leung2010}.  The MAPER whole brain segmentation algorithm
\citep{Heckemann2006,Heckemann2011}, using 30 atlases, on all ADNI1 baseline
images \citep{Heckemann2011}.  Lotjonen et al. use images from the ADNI1
baseline dataset and 30 atlases with a proprietary non-linear registration
method based on intensity differences, and post-processing step using a graph
cuts algorithm to optimise fused segmentations against a spatial intensity prior
\citep{Lotjonen2012}.  
\todo{No story here}

@@

\subsubsection{Automatic Normalization and Image Matching and Anatomical
Labeling (\animal)}

The \animal algorithm carries out image registration is in two phases. In the
first, a 12-parameter linear transformation (3 translations, rotations, scales,
shears) is estimated between images using an algorithm that maximizes the
correlation between blurred MR intensities and gradient magnitude over the whole
brain \citep{Collins}. In the second phase, nonlinear registration is completed
using the \animal algorithm \citep{Collins1995}: an iterative procedure that
estimates a 3D deformation field between two MR images. At first, large
deformations are estimated using blurred version of the input data. These larger
deformations are then input to subsequent steps where the fit is refined by
estimating smaller deformations on data blurred with a Gaussian kernel with a
smaller FWHM. The final transformation is a set of local translations defined on
a bed of equally spaced nodes that were estimated through the optimization of
the correlation coefficient.  For the purposes of this work we used the
regularization parameters optimized in Robbins et al. \citep{Robbins2004},
displayed in table \ref{tab:ANIMAL_params}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANIMAL parameters                                                           %
%                                                                             %
\begin{table*}[!tbp]
\scriptsize
\caption{\animal Registration Parameters\label{tab:ANIMAL_params}} 
\begin{center}
\begin{tabular}{l | c c c}
\hline 
Parameters        & Stage 1 & Stage 2 & Stage 3  \tabularnewline
\hline 
Model Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Input Blur (FWHM) &   8     &    8    &   4      \tabularnewline
Iterations        &   30    &    30   &   10     \tabularnewline
Step              &   8x8x8 &  4x4x4  &   2x2x2  \tabularnewline
Sub-Lattice       &   6     &    6    &   6      \tabularnewline
Lattice Diameter  &24x24x24 &12x12x12 & 6x6x6    \tabularnewline
\hline
\end{tabular}
\end{center}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Automatic Normalization Tools (\ants)}

\ants is a diffeomorphic registration algorithm which provides great flexibility
over the choice of transformation model, objective function, and the consistency
of the final transformation. The transformation is estimated in a hierarchical
fashion where the MRI data is subsampled, allowing large deformations to be
estimated and successively refined at later hierarchical stages (where the data
is subsampled to a finer grid). The deformation field and the objective function
are regularized with a Gaussian kernel at each level of the hierarchy. The \ants
algorithm is freely available \url{http://www.picsl.upenn.edu/ANTS/}. We used an
implementation of the \ants algorithm compatible with the MINC data format,
mincANTS \url{https://github.com/vfonov/mincANTS}.

We used the following command line when running \ants:
{\small
\begin{verbatim}
  mincANTS 3 -m PR[target_file.mnc,source_file.mnc,1,4] 
   --number-of-affine-iterations 10000x10000x10000x10000x10000 
   --affine-gradient-descent-option 0.5x0.95x1.e-4x1.e-4
   --use-Histogram-Matching --MI-option 32x16000
   -r Gauss[3,0] -t SyN[0.5] -i 100x100x100x20
   -o transformation.xfm
 \end{verbatim}
}
These settings were adapted from the "reasonable starting point" given in the
\ants manual 
\footnote{\url{https://sourceforge.net/projects/advants/files/Documentation/}}.

@@

The {\tt xcorr\_vol} utility from the \animal toolkit is used to calculate the
cross-correlation similarity measure.

@@

%%%%%
%%%%%                 Experiments: ADNI-1 cross-validationn               %%%%%
%%%%%
\subsubsection{ADNI1 cross-validation}
To test the effect of parameter settings on the \mb algorithm, Monte
Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a subset of
the ADNI1 dataset for validation.  This form of cross-validation (described in
detail below) allows us to validate \mb with various atlas and template
library sizes, registration and label fusion methods.

@@

{\bf Registration method.} Both the \ants and \animal registration methods were
used.

@@

A total of ten validation rounds are perfomed on each subject in the dataset,
for each combination of parameter settings. Reported subject segmentation
accuracy is the averaged over the ten validation rounds. The parameter settings
we explore are: atlas library size (1-9), template library size (1-20),
registration method (\ants vs.  \animal), and label fusion method (MV, XC-WV,
NMI-WV). A total of $10*69*9*20*2*3 = 7452000$ validation rounds were conducted,
resulting in a total of 1'490'400 segmentations analysed.

@@

{\bf Registration method.} The \ants non-linear registration algorithm was used,
as it performed best in the ADNI1 cross-validation experiment (see Results).

@@

{\bf Evaluation.}
Since the hippocampal segmentation protocols differ between the ADNI labels and
Winterburn atlases, this poses a problem for direct evaluation between labels
produced by \mb and the ADNI labels in terms of overlap; we would not expect
different segmentation protocols to have a high degree of overlap.  Instead, to
evaluate the performance of \mb we compare the correlation of \mb segmentation
volumes with manual segmentation (SNT) volumes.  Additionally, we correlate the
hippocampal volumes of established automated segmentation methods to \mb
segmentations.

@@

{\bf Registration method.} \ants.
\todo{complete}

@@

{\bf Registration method.} \ants
\todo{complete}

@@

%%%%%
%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
%%%%%
\subsection{ADNI1 Cross-Validation}

@@

In this experiment we conducted 10 rounds of \mb and multi-atlas segmentation of
each of 69 subjects at a range of atlas and template library sizes, registration
algorithm (\ants or \animal), and three label fusion techniques.  Hippocampal
\mb-based segmentations using both \animal and \ants registration algorithm
demonstrate good overlap with manually derived gold-standards (Figure
\ref{fig:ADNI1-xval-k-mean}). Qualitatively, both \animal and \ants-based
segmentations demonstrate trend overlap accuracy that increases with the size of
atlas library and template library. With 12 or more templates \mb library of at
least 12 templates

No marked difference in segmentation accuracy is seen when either \animal or \ants
registration is used with different label fusion techniques, at any atlas or
template library sizes.  In every parameter configuration, the use of \mb with
\ants registration shows increased segmentation accuracy over \mb with \animal
registration. In the remainder of this section, only results using the \ants
registration algorithm will be shown.

It is interesting to note that with an even number of templates, \mb shows a
small decrease in performance relative to when one fewer template image is used.
See section (\todo{ref}) for a discussion of this behaviour. In the remainder of
this section, only results from odd-sized template libraries will be shown.

@@

ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" )
@
  \label{fig:ADNI1-xval-k_mean}
  \caption{
  {\em \mb mean DSC by atlas and template library size, registration algorithm,
  and label fusion method. Mean DSC is calculated over each subject from the
  ADNI1 Cross-Validation } 
  }
\end{figure}

With an increasing number of templates, \mb-based segmentating using \ants
registration and majority vote label fusion shows improvement in overlap
accuracy over multi-atlas-based segmentation, using the same number of atlases
and voting method (Figure \todo{fig:ADNI1-xva-k-diff}. The magnitude of improvement over
multi-atlas-based segmentation decreases with an increasing number of atlases,
with accuracy converging with 7 atlases.  Peak improvement in \mb accuracy (~0.02
DSC) is found when one atlas is used with a template library of 20 images.
\todo{stats?}

\todo{mention failures}

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  coord_cartesian(ylim=c(-0.1,0.05)) + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean Kappa between MAGeT brain and multiatlas" ) 
@
  \caption{Increase in mean Kappa between \mb and multi-atlas}
  \label{ADNI1-xval-k-diff}
\end{figure}

@@

  #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
  #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
  #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
  #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
  labs(x="Number of Atlases", y = "Kappa", 
       title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
  scale_colour_hue(name="Method") + 
  coord_cartesian(ylim=c(0.7,0.9))
  #theme(legend.justification=c(1,0), legend.position=c(1,0))
@
  \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
  20 templates}
  \label{}
\end{figure}

@@

%%%%%
%%%%%                  Results: ADNI-1 Screening Validation               %%%%%
%%%%%
\subsection{ADNI1 Screening Validation}

@@

<<ADNI-src-volumes-prep,echo=F,cache=T>>=
bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Screening_1.5T_11_15_2012.csv")
#bl3T  = read.csv("data/ADNI_baseline_volumes/ADNI1_Baseline_3T_11_15_2012.csv")
#yr1   = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")
# recode VISCODE
#viscode_levels = list("bl"=1,"bl"=2, "m06"=3,"m12"=4)
#yr1$VISCODE <- as.factor(yr1$Visit)
#levels(yr1$VISCODE) <-viscode_levels
#bl1.5T$Scaled_2 <- grepl("Scaled_2", yr1$Description)

#subjects = subset(yr1,Visit ==1)
subjects = bl
subjects$VISCODE = 'bl'

# fetch just the columns we need, and do a little renaming
mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, 
                           MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, 
                           MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, 
                           SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
fs_pruned     = data.frame(TEMPQC =fs$TEMPQC, 
                           FS_L=fs$ST29SV, 
                           FS_R=fs$ST88SV, 
                           Source = fs$IMAGEUID)
fsl_pruned    = data.frame(FSL_L=fsl$X17, 
                           FSL_R=fsl$X53, Source=fsl$Source)

# diagnoses
dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])

# Now create one data.frame with all the yr1 volume data we have from each datasource
combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, dx          , by=c("RID","VISCODE"), all.x=TRUE)
combined = merge(combined, qc          , by=c("Image.Data.ID"), all.x=TRUE)

# A table showing the number of images (and failures) from each package
#           Total  MAGeT  FSL  FS   MAPER
# Images    
# Failures
# 
# Total number complete images =

# Step 2: Of those, get counts for the numbers that fail QC
package_totals = data.frame(
  Screening  = c(length(subjects$VISCODE), 'N/A'),
  SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
  MAGeT      = c(length(which(!is.na(combined$MAGeT_L)))  , length(which(combined$MAGeT_QC==0))), 
  MAPER      = c(length(which(!is.na(combined$MAPER_L)))  , "--"), 
  FSL        = c(length(which(!is.na(combined$FSL_L)))    , length(which(combined$FSL_QC==0))), 
  FS         = c(length(which(!is.na(combined$FS_L)))     , length(which(combined$TEMPQC == "Fail"))),
  row.names  = c("Images", "Failures")
)

all_qc = subset(all_qc, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )

attach(all_qc)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  #UCSD  = UCSD_L + UCSD_R,
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)
detach(all_qc)
totals.complete = na.omit(totals)
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 Screening Dataset Package Totals & Failures                           %
%                                                                             %
<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
latex(package_totals, file="", size="scriptsize", 
    caption="ADNI1 1.5T Screening Dataset Segmentations",
    title="",
    label="tab:ADN1-1.5T-Screening-Dataset-Segmentations")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
